Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory

### 2018-03-08 Cars/midcanPresentation.tex
presentation

  \documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

@@

\usepackage{alltt}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{alltt}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{tikz}
%\usetikzlibrary{bayesnet}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\setcounter{MaxMatrixCols}{30}

\def\app#1#2{%
  \mathrel{%
    \setbox0=\hbox{$#1\sim$}%
    \setbox2=\hbox{%
      \rlap{\hbox{$#1\propto$}}%
      \lower1.1\ht0\box0%
    }%
    \raise0.25\ht2\box2%
  }%
}
\def\approxprop{\mathpalette\app\relax}

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\setlength\parindent{0pt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\title{Online Updating of Variational Bayes for Heterogenous Forecasts of Vehicle Trajectory}
\author{Nathaniel Tomasetti}
\date{}

%TODO: Add a conclusion

\begin{document}
\maketitle

\section{Introduction}
\label{sec:intro}

Self-driving vehicles are rapidly becoming more advanced, with manufacturers such as Tesla, Ford, and Audi expecting autonomous vehicles to be available to consumers as early as 2020. These vehicles rely on a nagivation system that detects the position of surrounding traffic, forecasts their trajectery, and selects a path to avoid collisions. With as many as $94\%$ of accidents in the United States resulting from human error, according to the US National Highway Traffic Saftey Administration in 2015, \citep{NHTSA2015}, the introduction of fully self-driving vehicles is expected to substantially improve safety on the roads. However self-driving vehicles will operate in situations where the majority of surrounding vehicles are still controlled by human drivers, and so forecasts should be able to account for a wide range of possible human behaviour.
\\

While characteristics of surrounding vehicles, such as theirr postition, velocity, and steering angle, can be extracted from sensors through the use of deep neural networks (See e.g. \citet{Woo2016a} and \citet{Tian2017} for details), the problem of forecasting the future trajectory has seen less attention in the literature. These forecasts often assume the surrounding vehicles will maintain their current angle and velocity \citep{Gindele2010, Houenou2013, Bautista2017, Waymo2017} but interest has developed in more advanced models such as Neural Networks, Hidden Markov Models, and Support Vector Machines \citep{Ding2013, Woo2016b, Geng2017, Woo2017}.
\\

This paper takes a statistical approach to this trajectory forecasting problem, fitting time series models to two extracted variables, acceleration and steering angle, of vehicles in the Next Generation Simulation (NGSIM) US Highway 101 Dataset. It uses Bayesian methods to forecast the distribution of the future velocity and steering angle, which can be transformed into a forecast for the distribution of trajectories, allowing low probability trajectories that could cause a collision to be detected. The time series models used allow for heterogeneity between different vehicles, which may result from sources such as differences in individual driving styles, or the conditions of traffic. Including heterogeneity in the model will allow for drivers with extreme behaviours to be accounted for, for example 'lead-footed' drivers can be assigned a larger variance on their acceleration, while forecasts for more consistent drivers will benefit from a low variance.
\\

Incorporating heterogeneity into forecasts requires the behaviour of each driver of interest to be inferred as vehicles are encountered while driving. This inference is facilitated by first obtaining the posterior distribution for a large number of vehicles before the self-driving vehicle is on the road, from which the prior distribution for additional vehicles can be constructed. Once the self-driving vehicle is on the road, online inference is made possible by the introduction of Updating Variational Bayes (UVB), which approximates Bayesian updating so that inference may be periodically updated as additional data relating to surrounding vehicles is made available. 
\\

We find that the use of time series models produces accurate point forecasts in terms of mean Euclidean error when compared to the contstant models commonly used, and including driver heterogeneity in these time series models substantially reduces the uncertainty in the entire forecast distribution relative to a homogeneous approach. These results give credence to the validity of the methodological framework developed in this thesis, where exact Bayesian inference techniques are used to guide the construction of a prior and approximating distribution for parameters that are inferred in an online setting with Variational Bayes.
\\

This paper is structured as follows: Section \ref{sec:dataProcessing} describes the dataset used and steps required to extract acceleration and steering angle for each vehicle, while Section \ref{sec:models} introduces several time series models used. Section \ref{sec:Inference} reviews Bayesian inference, which is extended to the online, approximate updating setting in Section \ref{sec:UVB}. Finally Section \ref{sec:eval} contains empirical results, and Section \ref{sec:disc} concludes the vehicle trajectory forecasting study and discusses future work.

\section{Data processing}
\label{sec:dataProcessing}
Data is provided by the Next Generation Simulation (NGSIM) project conducted by the US Federal Highway Administration (FHWA), which recorded vehicles travelling along a 2235 feet section of the US 101 freeway in Los Angeles, California during the morning peak from 7:50 am to 8:35 am on June 15th, 2005. Data for 6101 vehicles was collected every 100 milliseconds by seven static cameras, and processed by Cambridge Systematics Inc to produce coordinates for each vehicle and point in time relative to the start of the road section. 
Figure \ref{fig:rawData} shows the paths of ten vehicles in the dataset, which traveled along the five major lanes or entered from the entry/exit lane to the right; with one vehicle changing from Lane 2 to Lane 1. Vehicles that entered or exited midway through the freeway section are excluded in this paper. There is a curvature to the road occuring between 500 and 1000 feet, and again between 1800 and 2000 feet.
\\

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{carPath}
\caption{The path of ten vehicles in the dataset, with each black line representing a unique vehichle. This section of US 101 is split into five main lanes, with an additional entry/exit lane to the far right, through which two of the vehicles have entered. There is curvature to the road (which is slightly distorted by the aspect ratio) with bends occuring between 500 and 1000 feet, and again between 1800 and 2000 feet.}
\label{fig:rawData}
\end{figure}

Modern vehicles are capable of tracking the path of lane markings \citep{Thuy2010}, such as painted centre lines or lane dividers, and thus they can automatically identify the locations of surrounding vehicles in coordinates relative to their own position on the road compensating for any curvature. To remove the variation in car position due to curvature in the road, and so that the data is in a coordinate system similar to that of a self-driving vehicle tracking surrounding vehicles, the coordinate system provided by NGSIM is transformed into relative coordinates. This is facilitated by measuring the location of vehicles relative to an estimate of the midpoint of their associated lane. \citet{Woo2016a} fit a polynomial curve to detected lane markings to build a local model of the road lane edges, and this idea is extended with the use of smoothing splines to estimate each of the lane centres for the available section of the freeway. Details of this process are in the appendix, and results in relative coordinates $\{x^*_{i, t}, y^*_{i, t}\}$  where $y^*_{i, t}$ denotes the distance travelled along the cente of the road, and $x^*_{i, t}$ denotes the deviation from the lane centre line for each of remaining vehicles in the dataset.
\\

Once relative coordinates are extracted, the changes in relative position of any vehicle $i$ from observation $t-1$ to observation $t$ follows friom the trigonometric relationship in Figure \ref{fig:motion}:
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq} \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq},
\end{align}
Note that when $\delta_{i, t} \pm \pi/2$ then car $i$ takes a position parallel to the centre line, and hence $x^*_{i, t} = x^*_{i, t-1}$.
\\

From this relationship, and the coordinate sequences $\{x^*_{i, s}, \mbox{ for }s=1,2,...,T\}$ and $\{y^*_{i, s}, s=1,2,...,T\}$, the inputs to motion from driver $i$ are calculated via
\begin{align}
\delta_{i, t} &= 
     \begin{cases}
       \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right)  &\quad\text{if }x^*_{i, t} \neq x^*_{i, t-1} \\
       \frac{\pi}{2} &\quad\text{if } y^*_{i, t} > y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       -\frac{\pi}{2} &\quad\text{if } y^*_{i, t} < y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       \delta_{i, t-1} &\quad\text{otherwise,} \\ 
     \end{cases} \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}.
\end{align}
In addition, the corresponding acceleration sequence is given by $\{a_{i, t},t=3,4,...,T\}$, according to
\begin{equation}
\label{aEq}
a_{i, t} = v_{i, t} - v_{i, t-1}. 
\end{equation}
\\

Data resulting from the observation of vehicle $i$ for a period of time $T$ is defined as $z_{i, 1:T} = \{x^*_{i, s}, y^*_{i, s}, | s = 1, \dots, T\}$ and data from multiple drivers are then collected as $\mathbf{z}_{1:N} = \{z_{i, 1:T} | i = 1, \dots, N\}$. Some additional vehicles are excluded due to an insufficient number of observations, leaving 3387 vehicles in the dataset in a relative coordinate system. These vehicles are split into two sets of vehicles: a training set of 2000 vehicles and a test set of a further 1387 vehicles.
\\

The training set and the test set are treated as observations from vehicles at two distinct time periods, illustrated in Figure \ref{fig:problem}. The training set in the left panel represents any vehicle for which data is collected during the development of the self-driving vehicle, for example this data could be collected by human driven vehicles. The test set in the right panel represents those vehicles the self-driving vehicle will encounter when it is driving autonomously, which require trajectory forecasts in close to real time.

\begin{figure}
\centering
\includegraphics[width = 0.4\textwidth]{motion}
\caption{A vehicle at coordinate $\{x^*_{t-1}, y^*_{t-1}\}$ at time $t-1$ will be at $\{x^*_t, y^*_t\}$ at time $t$ if its angle over this period is $\delta_t$ and velocity is $v_t$.}
\label{fig:motion}
\end{figure}

\begin{figure}
\centering
\includegraphics[width = 0.95\textwidth]{problem}
\caption{Left: The paths of vehicles in the training set. Right: The self-driving car (Solid) and two nearby vehicles (Outlined). All motion is from left to right. The self-driving vehicle must forecast the trajectory of the nearby vehicels based on their observed path and any knowledge of the driver behaviour learned from the training set vehicles}
\label{fig:problem}
\end{figure}

\section{Models for Trajectory Forecasting}
\label{sec:models}

Three related time series models are introduced to produce trajectory forecasts, differing by their implications about the dependency between the different sets of vehicles. These are referred to as the homogenous model, under which all drivers are treated identically and the training set is used to infer all behaviour and two heterogenous models: the independent heterogenous (IH) model and the hierarchical, clustered heterogeneity (CH) model. The ID model does not incorporate information from the vehicles in the training into inference for the test set, while the CH model allows information to be shared between vehicles; learning a range of possible behaviours from the training set that can be incorporated into inference about the test set. 
\\

\subsection{A Homogenous Time Series Model}
\label{subsec:homogenous}

The framework provided by (\ref{xEq}) and (\ref{yEq}) allows distributional forecasts of future position coordinates of a given vehicle $i$, $\{x^*_{i, t}, y^*_{i, t} | t = T + 1, \dots, T+H\}$, to be obtained given the position and velocity of a vehicle at time $T$, $\{x^*_{i, T}, y^*_{i, T}, v_{i, T}\}$, and distributional forecasts of the future values of $\{a_{i, t}, \delta_{i, t} | t = T + 1, \dots, T+H\}$.  
\\

\iffalse % Partial Autocorrelation plots are a better way to show dynamic behaviour in the data.
This is highlighted in Figure \ref{fig:dynamics}, where the left panel displays a plot of the mean value of
\begin{equation}
A_{i, s} = \frac{a_{i, \tau_i + s}}{a_{i, \tau_i}}
\label{amax}
\end{equation}
where $\tau_i = \arg \underset{t}{\max}|a_{i, t}|$, the relative time $s$ takes on values between -1000 and 500 milliseconds, and with the average taken over the $i = 1, 2, \dots, N$ individual cars. On average, large vehicle acceleration or deceleration is preceeded by an increase in acceleration in the same direction for 400 milliseconds. The right panel displays a plot of the mean value of
\begin{equation}
D_{i, s} = \frac{\delta_{i, \omega_i + s} - \pi/2}{\delta_{i, \omega_i} - \pi/2}
\label{dmax}
\end{equation}
where $\omega_i = \arg \underset{t}{\max}|\delta_{i, t} - \pi/2|$, for again $-1000 \leq s \leq 500.$ Simmilarly, this second panel demonstrates that the largest steering angle deviation from $\pi/2$ may be anticipated by smaller deviations in the same direction, at typically, some point during the previous 300 milliseconds. These plots sugget that changes in acceleration and angle are predictable if given the recent history of a vehicle's position. 
\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth]{dynamics}
\caption{Left: Mean values of acceleration as a percent of a vehicles maximum absolute acceleration for the 1000 milliseconds before, and 500 milliseconds after, a vehicle reaches its maximum absolute acceleration. Right: The analogous plot for the steering angle, where deviations in angle from $\pi/2$ are plotted as a percentage of a vehicles maximum absolute value of deviation. The few observations before either variable reaches its maximum often have increased values in the same direction.}
\label{fig:dynamics}
\end{figure}
\fi

NGSIM vehicle data is observed every 100 milliseconds so large changes in $a_{i, t}$ and $\delta_{i, t}$ may occur over multiple sequential observations. This is highlighted in Figure \ref{fig:pacf}, which displays a plot of the partial autocorrelation function, which measures the correlation between a variable at two different time periods, $t$ and $t + k$, conditioned on that variable at time $t + 1, \dots t + k -1$, for both $a_{i, t}$ and $\delta_{i, t}$ for five vehicles. Large spikes in the partial autocorrelation plots, such as those at the first and second lag for most vehicles, demonstrate that the value of either $a_{i, t}$ or $\delta_{i, t}$ is strongly correlated with the previous time periods. 
\\

\begin{figure}
\centering
\includegraphics[width = 0.95\textwidth]{pacf}
\caption{Top: Partial Autocorrelation plots for acceleration, $a$, for a random selection of vehicles. Bottom: Partial Autocorrelation plots for steering angle, $\delta$ for the same vehicles. Each spike for vehicle $i$ at lag $k$ in the top panel indicates that $a_{i, t-k}$ is correlated with $a_{i, t}$, similarly spikes in the bottom panel indicate that $\delta_{i, t-k}$ is correlated with $\delta_{i, t}$. Note that different vehicles do not have the same dynamic behaviour in the partial autocorrelations.}
\label{fig:pacf}
\end{figure}

>>>>

\documentclass[8pt]{beamer}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

@@

\usepackage{alltt}

\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}

\usepackage{alltt}
\usepackage{natbib}
\usepackage{array}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage{bm}
\setbeamertemplate{caption}[numbered]

\usetheme{Warsaw}
\defbeamertemplate*{footline}{shadow theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fil,rightskip=.3cm]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertframenumber\,/\,\inserttotalframenumber\hfill\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle%
  \end{beamercolorbox}}%
  \vskip0pt%
}

\setbeamertemplate{headline}{%
\leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=2.5ex,dp=1.125ex]{palette quaternary}%
    \insertsectionnavigationhorizontal{\paperwidth}{\hskip0pt plus1filll}{\hskip0pt plus1filll}
    \end{beamercolorbox}%
  }
}

\newcounter{saveenumi}
\newcommand{\seti}{\setcounter{saveenumi}{\value{enumi}}}
\newcommand{\conti}{\setcounter{enumi}{\value{saveenumi}}}
\resetcounteronoverlays{saveenumi}

\title[Online Updating of Variational Bayes]{Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\begin{frame}
\titlepage
\centering
Supervised by Catherine Forbes and Anastasios Panagiotelis
\end{frame}

\begin{frame}
\tableofcontents
\end{frame}

\begin{frame}
\section{Framework}
\frametitle{Thesis Framework}
Focus: Two sets of data
\begin{itemize}
\item Stuff Here about the general themes
\item Use of exact inference to improve VB
\item Hierarchical models with exact inference
\item Some new unit with time constrained inference
\item Want to use MCMC info to improve VB
\end{itemize}
\end{frame}

\begin{frame}
\section{Inference}
\frametitle{Bayesian Inference}
Note: reframe discussion with two sets of data from previous slide
\begin{itemize}
\item The posterior distribution for parameters $\theta$, $p(\theta | \textbf{z})$, can be inferred with Markov Chain Monte Carlo (MCMC). MCMC:
\begin{enumerate}
\item Allows 'exact' inference as the number of iterations increases
\item Can be very slow
\end{enumerate}
\item We focus on settings where MCMC inference for the additional unit is too slow to be useful.
\item Instead, Variational Bayes (VB) is used. VB:
\begin{enumerate}
\item Approximates the posterior with an analytical alternative $q_{\lambda}(\theta | \textbf{z})$.
\item Chooses the distribution $q$ that minimises a divergence funciton
\item Is typically much faster than MCMC.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Kullback Leibler Divergence and Stochastic Optimisation}
\begin{itemize}
\item The typical divergence is the Kullback Leibler Divergence from $q$ to $p$.
\end{itemize}
\begin{equation}
\label{KL-def}
KL[q_{\lambda}(\theta | \textbf{z})\hspace{.1cm}||\hspace{.1cm}p(\theta | \textbf{z})] = E_{q_{\lambda}(\theta | \textbf{z})} \left[ \log(q_{\lambda}(\theta | \textbf{z})) - \log(p(\theta | \textbf{z})) \right],
\end{equation}
\begin{itemize}
\item Monte-Carlo simulations are usually unavailable due to the $p(\theta | \textbf{z})$ term.
\end{itemize}
\begin{equation}
\label{ELBO}
\mathcal{L}(q, \lambda) = E_{q_{\lambda}(\theta | \textbf{z})} \left[\log(p(\theta, \textbf{z})) - \log(q_{\lambda}(\theta | \textbf{z}))\right],
\end{equation}
\begin{itemize}
\item Stochastic Estimates are available through Monte-Carlo:
\end{itemize}
\begin{equation}
\label{ELBO-MC}
\mathcal{L}(q, \lambda) \approx \frac{1}{M} \sum_{j=1}^M \left(\log(p(\theta_{j}, \textbf{z})) - \log(q_{\lambda}(\theta_{j} | \textbf{z})) \right)
\end{equation}
where $\theta_{j} \sim q_{\lambda}(\theta | \textbf{z})$.
\end{frame}

\begin{frame}
\frametitle{Gradient Estimators}
\begin{itemize}
\item For most $p$ and $q$ only Stochastic Gradient Ascent is available for optimisation. Repeatedly apply:
\begin{equation}
\label{gradientAscent}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}} \bigg\rvert_{\lambda = \lambda^{(m)}}
\end{equation}
\item ELBO gradients can be estimated by the Score Estimator:
\begin{equation}
\label{scoreDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{SC} = \sum_{j = 1}^M \frac{\delta \log(q_{\lambda}(\theta_{j, i} | \textbf{z}))}{\delta \lambda} \left(\log(p(\theta_{j, i}, \textbf{z})) - \log(q_{\lambda}(\theta_{j, i} | \textbf{z})) \right),
\end{equation}
where $\theta_j \sim q_{\lambda}(\theta | \epsilon)$.
\item Or the Reparametersied Estimator, where $\theta = f(\epsilon, \lambda)$:
\begin{equation}
\label{rpDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{RP} = \sum_{j = 1}^M \frac{\delta \theta}{\delta \lambda} \frac{\delta \log(p(\theta, \textbf{z}))}{\delta \theta} \bigg\rvert_{\theta = f(\lambda, \epsilon_j)} + \frac{\delta J(\lambda, \epsilon_j)}{\delta \lambda}, 
\end{equation}
where $\epsilon_j \sim q(\epsilon)$.
\end{itemize}
\end{frame}

\begin{frame}
\section{Online Heterogenous Forecasting: Problem}
\subsection{Motivation}
\frametitle{Self-Driving Vehicles}
\begin{itemize}
\item Self-Driving Vehicles are a thing soon
\item The navigation system has three steps
\begin{enumerate}
\item Detect surrounding traffic
\item Forecast surrounding traffic
\item Navigate through safely
\end{enumerate}
\item Forecasting has not had much work
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Aims of the first paper}
\begin{itemize}
\item What we plan on doing with this first paper
\item Overall goals and stuff
\item Motivate heterogeneity
\item Note heterogeneity needs inference on vehicles as they are encountered
\item There is not much time to do that
\item Develop an approximation to the Bayesian Update Equation so we can infer parameters for new vehicles online
\item Also mention something on density forecasting would be nice
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Data}
\frametitle{Next Generation Simulation}
\begin{itemize}
\item This is a dataset created for macro traffic modelling
\item It records all of the cars all of the time
\item The road is curved
\item Before and After of road straightening project
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Forecasting in the framework}
\begin{itemize}
\item That is a terrible title.
\item Split the data into two sets: train/test
\item Cars observed beforehand, cars encountered while driving
\end{itemize}
\begin{figure}
\centering
\includegraphics[width = 0.95\textwidth]{problem}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Trigonometric Motion Equations}
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq}, \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq}.
\end{align}
Picture
\end{frame}

\begin{frame}
\frametitle{Extracting Driver Inputs}
\begin{align}
\delta_{i, t} &= 
     \begin{cases}
       \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right)  &\quad\text{if }x^*_{i, t} \neq x^*_{i, t-1} \\
       \frac{\pi}{2} &\quad\text{if } y^*_{i, t} > y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       -\frac{\pi}{2} &\quad\text{if } y^*_{i, t} < y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       \delta_{i, t-1} &\quad\text{otherwise,} \\ 
     \end{cases} \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}.
\end{align}
\begin{equation}
\label{aEq}
a_{i, t} = v_{i, t} - v_{i, t-1}. 
\end{equation}
\begin{itemize}
\item Why other variables are useful
\item Acceleration is good because relative velocity / non-stationarity
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Modelling}
\frametitle{Auto-regressive Models for Trajectory Forecasting}
\begin{itemize}
\item Evidence of dynamics behaviour
\item Modelling approachs / relationship to test train set
\item Motivate the use of the three different models
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A Homogeneous Approach}
Talk about what is going on here
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{j} a_{i, t-j} + \sigma_{\epsilon} \epsilon_{i, t} \label{aAR} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta} \eta_{i, t} \label{dAR}
\end{align}
\begin{equation*}
\label{thetaVec}
\theta = \{\phi_{1}, \dots, \phi_{p}, \gamma_{1}, \dots, \gamma_{q}, \log(\sigma^{2}_{\epsilon}), \log(\sigma^{2}_{\eta})\}.
\end{equation*}
\begin{equation}
\label{indPrior}
\theta \sim \mathcal{N}\left(\mu, \Sigma \right)
\end{equation}
\begin{itemize}
\item The Posterior
\item Why are we doing this
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Independent Heterogeneity Model}
Maybe different drivers behave differently?
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{i, j} a_{i, t-j} + \sigma_{\epsilon, i} \epsilon_{i, t} \label{aAR2} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{i, j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta, i} \eta_{i, t}. \label{dAR2}
\end{align}
This breaks down into a bunch of independent posteriors
\begin{equation}
p(\theta_{1:N} \mid z_{1:N}) = \prod_{i=1}^N p(\theta_{i} \mid z_{i}),
\end{equation}
The training set isn't terribly useful.
\begin{equation}
p(\theta_{i}| \textbf{z}_{1:N}, z_{i,1:T}, \theta_{1:N}) = p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_i) p(\theta_i).
\label{indepNewCar}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
Hierarchical Models share information 
\begin{itemize}
\item Introduce Auxiliary Variable K
\item Sharing of the information
\end{itemize}
\begin{equation}
\label{mixPrior}
\theta_i | k_i = j \sim N(\mu_j, \Sigma_j),
\end{equation}
and
\begin{equation}
k_i \mid \beta \sim \mbox{Multinomial}\left(\pi_1, \dots, \pi_{K}\right)
\end{equation}
\begin{align}
\mu_j &\sim \mathcal{N}\left(\bar{\mu}_j, \Omega_j\right), \\
\Sigma_j &\sim \mbox{Inverse Wishart}\left(\mbox{Degrees of Freedom } \tau_j, \mbox{Scale } \Psi_j\right), \\
\boldsymbol{\pi} &\sim \mbox{Dirichlet}\left(\alpha_1 = \alpha_2 = \dots = \alpha_K\right).
\end{align}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
The new vehicle depends on the old vehicles through $\beta$.
\begin{equation}
\label{hierNewCar}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) \int_{\beta} p(\theta_{i} | \beta) p (\beta | \textbf{z}_{1:N}) d\beta.
\end{equation}
The $\beta$ posterior is precise for large $N$
\begin{equation}
\label{hierNewCar2}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) p(\theta_{i} | \hat{\beta}).
\end{equation}
\end{frame}

\begin{frame}
\begin{itemize}
\item That picture showing the IH / CH fits
\item Evidence of heterogeneity and the suitability of the mixture model
\end{itemize}
\end{frame}

\begin{frame}
\section{Updating Variational Bayes}
\frametitle{Updating Variational Bayes}
\begin{itemize}
\item Motivation
\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth]{timeUpdate}
\end{figure}
\item Notation: $q_{\lambda_S}(\theta_{i} | z_{i, 1:S})$ and $q_{\lambda_T}(\theta_{i} | z_{i, 1:T})$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Approximating Bayesian Updating}
\begin{equation}
\label{updatePost}
p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S})
\end{equation}
\begin{itemize}
\item This only works if the right hand side can be evaluated
\item VB Gradient Estimators requires the joint distribution
\begin{equation}
\label{updateJoint}
p(\theta_{i}, z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S})
\end{equation}
\item We will replace it with this
\begin{equation}
\label{ApproxJoint}
\hat{p}(\theta_{i},  z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})q_{\lambda_S}(\theta_{i} | z_{i, 1:S}).
\end{equation}
\item Sources of error: How close $q$ is to $p$
\end{itemize}
\end{frame}
